Resilience and Home-Space for WSTS \thanks{This work was partly done while the authors were supported by the Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028).}

### 2023-10-31 VMCAI-revised-version/final.tex
Appendix minsk back again

\section{Structural halting in reset-VASS}

We claim {\sc Structual existential halting} in reset-VASS is undecidable.  
This is based on the construction providng the undecidability of structural termination in Lossy counter machines and in reset-VASS~\cite{dufourd1998reset,mayr1998lossy}. Strictly speaking undecidability of
{\sc Structual existential halting} is not considered in
\cite{mayr1998lossy},
but 
the encoding of Minsky machines can be reused
since 
 {\sc Structual existential halting}
 is equivalent to
 structural termination in the case of deterministic Minsky machines and is thus
 undecidable for Minsky machines in general.

>>>>

\section{Reset-VASS and Minsky-machines}\label{minsk}

We now delve into more details into the undecidability results 
upon which rely
Proposition~\ref{indec WSTS with dcs}
and
Proposition~\ref{srp down up},
namely, that
{\sc Resilience} and {\sc State-resilience} when $\Safe = \mathop{\downarrow}~\Safe$
are undecidable for Minsky machines
and 
for reset-VASS.

The problems are undecidable for a Minsky machine with at least $2$ counters when
$\Safe$ is downward-closed, since it is possible to consider $\Safe = \{q_f(0,0)\}$. We then use Reset-VASS to simulate the Minsky machines~\cite{araki1976PN,dufourd1998reset}, albeit in a weak way. 
% The simulation however is 
Since by a particular choice of $\Safe$ and $\Bad$ we are only interested in the reachability of $q_f(0,0)$ in the Minsky machine, 
this weak simulation will be enough to obtain undecidability of both problems.

\begin{proposition}\label{both minsk indec}
{\sc Resilience} and {\sc State-resilience} when $\Safe = \mathop{\downarrow}~\Safe$ are undecidable for reset-VASS, hence also undecidable for $\post$-effective, $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility.
\end{proposition}

\begin{proof}
First, let us show that {\sc Resilience} is undecidable for Minsky machine with at least $2$ counters.
	This is due to the undecidability of $2$-counter Minsky machine termination.
From a $2$-counter Minsky machine $M$ with final control-state $q_f \in Q_M$, one can construct a $2$-counter Minsky machine $M'$ with final control-state $q_f' \in Q_{M'}$
such that there exists a terminating run in $M$ for all inputs iff machine $M'$ can reach $q_f'(0,0)$ from any input. 
%			with at least $1$ on its third counter. 
We build $M'$ to simulate $M$ until it reaches $q_f$, then decrease the two counters until they reach $0$, then, and only then, go in $q_f'$.
Based on this construction, the problem of deciding whether the downward-closed set 
$\{q_f'(0,0)\}$ is reachable in the $2$-counter Minsky machine $M'$
from any input is equivalent to the asking whether there exists a terminating run from every input of machine $M$, that is undecidable~\cite{Min61}. 
Note that by chosing $s_0 \in Q_{M'} \times \N^2 \setminus \{q_f'(0,0) \}$,
deciding whether the downward-closed set 
$ \{q_f'(0,0) \}$ is reachable
from every element of $\post^*(s_0)$ 
is again equivalent to asking whether $M$ terminate on every input.
Hence {\sc Resilience} and {\sc State-resilience} are undecidable for Minsky machines.

% \alain{écrire qqchose ici}
Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset}. 
% \alain{\cite{araki1976PN} simule-t-il les Minsky machines et comment ?}
%Reset-VASS extend the basic VASS model with special “reset
%transitions” that set to $0$ some coordinates in the vector. 
%			Let us recall their definition here.
%      	it is well known that reset-VASS are WSTS~\cite{dufourd1998reset}.
The simulation operates by building a reset-VASS in which
there exists
 a sequence from 
an initial state %$(s,x_1 \ldots x_n)$
to
a state $ q({\bf 0})$ if and only if
there exists
a sequence
 in the Minsky machine
 from a corresponding initial state
 to
 the state $q_f(0,0,0)$
  (from Theorem~$5$ in \cite{araki1976PN}).
Hence {\sc Resilience} (resp. {\sc State resilience}) holds in the Minsky machine $M'$ iff 
{\sc Resilience} (resp. {\sc State resilience}) holds in the reset-VASS simulating $M'$.
Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} 
and  {\sc State-resilience} are undecidable for reset-VASS and hence for  $\pred$-basis-effective, 
$\omega^2$-WSTS with strong compatibility as well.\qed
\end{proof}

\iffalse
\section{Structural halting in reset-VASS}

We claim {\sc Structual existential halting} in reset-VASS is undecidable.  
This is based on the construction providng the undecidability of structural termination in Lossy counter machines and in reset-VASS~\cite{dufourd1998reset,mayr1998lossy}. Strictly speaking undecidability of
{\sc Structual existential halting} is not considered in
\cite{mayr1998lossy},
but 
the encoding of Minsky machines can be reused
since 
 {\sc Structual existential halting}
 is equivalent to
 structural termination in the case of deterministic Minsky machines and is thus
 undecidable for Minsky machines in general.
\fi
